\input{../style.tex}

\title{Putnum and Beyond Solution Manual}
\author{Jacob Denson}

\begin{document}

\pagenumbering{gobble}
\maketitle
\tableofcontents
\pagenumbering{arabic}

\chapter{Basic Concepts}

\section{Preliminaries}

\begin{exercise}
    Basic Vector Space Terminology.
    \begin{enumerate}
        \item[(a)] Show that if $A$ is an absorbing set or a nonempty balanced set, then $0 \in A$.
        \begin{proof}
            If $A$ is absorbing, there is $\lambda > 0$ for which $0 \in \lambda A$. But then
            %
            \[ 0 = \lambda^{-1} 0 \in \lambda^{-1} \lambda A = A \]
            %
            If $A$ is a non-empty balanced set, then $0 \in 0 A \subset A$, since $|0| < |1|$.
        \end{proof}

        \item[(b)] Show that if $A$ is balanced, then $\alpha A = A$ whenever $|\alpha| = 1$.
        \begin{proof}
            It is obvious that $\alpha A \subset A$. Conversely, since $|\alpha^{-1}| = 1$, $\alpha^{-1} A \subset A$. Given $a \in A$, $\alpha^{-1} a \in \alpha^{-1} A \subset A$, but then $a = \alpha(\alpha^{-1} a) \in \alpha A$.
        \end{proof}

        \item[(c)] Suppose that $\mathcal{B}$ is a collection of balanced subsets of $X$. Show that $\bigcup \{ S : S \in \mathcal{B} \}$ and $\bigcap \{ S : S \in \mathcal{B} \}$ are both balanced.
        \begin{proof}
            For any $|\alpha| \leq 1$, $B \in \mathcal{B}$, $\alpha B \subset B$, so that
            %
            \[ \alpha \bigcap_{B \in \mathcal{B}} B = \bigcap_{B \in \mathcal{B}} \alpha B \subset \bigcap_{B \in \mathcal{B}} B \]
            %
            \[ \alpha \bigcup_{B \in \mathcal{B}} B = \bigcup_{B \in \mathcal{B}} \alpha B \subset \bigcup_{B \in \mathcal{B}} B \]
            %
            and therefore the union and intersection of balanced sets is balanced.
        \end{proof}

        \item[(d)] Suppose that $\mathcal{C}$ is a collection of convex subsets of $X$. Show that $\bigcap \{ S : S \in \mathcal{C} \}$ is convex.
        \begin{proof}
            If $C \in \mathcal{C}$ $a, b \in C$, $\lambda \in [0,1]$, then $\lambda a + (1 - \lambda) B \in C$. By putting $\forall C \in \mathcal{C}$ in the front of these statements, we obtain the statement for the intersection.
        \end{proof}

        \item[(e)] Show that if $A$ is convex, then $x + A$ and $\alpha A$ are convex.
        \begin{proof}
            If $x + a, x + b \in x + A$, $\lambda \in [0,1]$, then
            %
            \[ \lambda (x + a) + (1 - \lambda) (x + b) = x + (\lambda a + (1 - \lambda) b) \in x + A \]
            %
            and therefore $x + A$ is convex.
        \end{proof}
    \end{enumerate}
\end{exercise}

\begin{exercise}
    \begin{enumerate}
        \item[(a)] Show that the ``addition'' and ``multiplication by scalars'' defined for sets obey the commutative and associative laws for vector spaces. That is, show that $A + B = B + A$, that $A + (B + C) = (A + B) + C$, and that $\alpha (\beta A) = (\alpha \beta) A$. Show also that $(x + A) + (y + B) = (x + y) + (A + B)$.
        \item[(b)] Show that $\alpha(A + B) = \alpha A + \alpha B$.
        \item[(c)] Show that $(\alpha + \beta)A \subset \alpha A + \beta A$, but that equality does not always hold.
        \begin{proof}
            The equations can be verified pointwise. If the equations is satisfied on the left side by a point, it holds on the right side, and vice versa. This is not true of the third question, since 
        \end{proof}
    \end{enumerate}
\end{exercise}

\begin{exercise}
    \begin{enumerate}
        \item[(a)] Prove that $A$ is convex if and only if $sA + tA = (s + t)A$ for all positive $s$ and $t$. (Consider the special case in which $s + t = 1$).
        \begin{proof}
            s
        \end{proof}
    \end{enumerate}
\end{exercise}

\section{Norms}

\section{First Properties of Norm Spaces}

\begin{exercise}
    Let $K$ be a compact Hausdorff space and let $X$ be a normed space. By Corollary 1.3.4, the collection of all continuous functions from $K$ into $X$ is a vector spacewhen functions are added and multiplied by scalars in the usual way. Define a norm on this vector space by the formula
    %
    \[ \| f \|_\infty = \begin{cases} \max \{ \|f(x)\| : x \in K \} & \text{if}\ K \neq \emptyset \\ 0 & K = \emptyset \end{cases} \]
    %
    The resulting normed space is denoted $C(K,X)$.
    %
    \begin{enumerate}
        \item[(a)] Show that $\| \cdot \|_\infty$ is in fact a norm on $C(K,X)$.
        \begin{proof}
            $\| f + g \| \leq \| f \| + \| g \|$, since $\sup (A + B) \leq \sup A + \sup B$ for any $A$ and $B$. $\| \alpha f \| = |\alpha| \| f \|$, since
            %
            \[ \max \{ \|\alpha f(x)\| : x \in K \} = |\alpha| \max \{ \|f(x)\| : x \in K \} \]
            %
            And if $\| f \| = 0$, then $\|f(x)\| = 0$ for all $x$, so that $f(x) = 0$ for all $x$.
        \end{proof}

        \item[(b)] Show that if $X$ is a Banach space, then so is $C(K,X)$.
        \begin{proof}
            Let $f_1, f_2, \dots$ be a Cauchy sequence in $C(K,X)$, so that $\| f_i - f_j \| \to 0$.
        \end{proof}
    \end{enumerate}
\end{exercise}

\end{document}