\input{../../style.tex}

\title{Graph Theory}
\author{Jacob Denson}

\begin{document}

\pagenumbering{gobble}

\maketitle

\tableofcontents

\pagenumbering{arabic}

\chapter{Basic Graph Theory}

Many situations can be formalized by diagrams involving points connected between one another, in such a way that the particular way they are connected is immaterial. Graph theory is the mathematical study of such situations.

\section{Bipartite Graphs}

A graph $G$ is \emph{bipartite} if it's vertices can be divided into two disjoint sets $V_1$ and $V_2$ such that every edge in $G$ has one end in $V_1$ and one end in $V_2$. Given such a decomposition, we say elements of $V_1$ are `left hand vertices', and elements of $V_2$ are `right hand vertices'.

\begin{theorem}
	A graph $G$ is bipartite if and only if every cycle in $G$ has even length.
\end{theorem}
\begin{proof}
	One direction is clear, and so it suffices to show that if every cycle in $G$ has even length has even length, then $G$ is bipartite. We proceed by induction, the base case being the trivial graph with no vertices, which is obviously bipartite. Without loss of generality, we may assume that $G$ is connected. For the inductive case, deleting some vertex $v$ to form a graph $G'$ with connected components $H_1 \cup \dots \cup H_n$. By induction, each graph $H_i$ is bipartite, so we can write the vertices of $H_i$ as a disjoint union $V_{i1} \cup V_{i2}$ which gives the graph a bipartite structure. We claim that for each $i \in \{ 1, \dots, n \}$, there exists $j_i \in \{ 1, 2 \}$ such that $v$ is only connected to vertices in $V_{i j_i}$, because otherwise we could find an odd length cycle containing $v$. If we write $V_1 = V_{1j_1} \cup \dots \cup V_{1j_n}$ and $V_2 = V_{1j_1}^c \cup \dots \cup V_{1j_n}^c \cup \{ v \}$, then this gives a bipartite structure on $G$.
\end{proof}

\begin{theorem}
	Let $G$ be a bipartite graph with vertex set $V_1 \cup V_2$, where $V_2$ is nonempty, and edge set $E$. Then $G$ has a bipartite subgraph $G'$ with vertex set $V_1' \cup V_2'$ and edge set $E'$, where $\#(E') \geq \#(E)/2$, and $\deg_{G'}(v_2) \geq \#(E)/2\#(V_2)$ for each $v_2 \in V_2'$.
\end{theorem}
\begin{proof}
	For any integer $N$,
	%
	\[ \sum_{\deg(v_2) > N} \deg(v_2) \geq \#(E) - N \#(V_2). \]	
	%
	Thus if $N = \#(E) / 2 \#(V_2)$,
	%
	\[ \sum_{\deg(v_2) > N} \deg(v_2) \geq \#(E)/2, \]
	%
	so we can set $V_1' = V_1$, and $V_2' = \{ v_2 \in V_2: \deg(v_2) > \#(E)/2\#(V_2) \}$, and then let $E'$ be the set of all vertices from $V_1'$ to $V_2'$.
\end{proof}

\begin{theorem}
	Let $G$ be a bipartite graph with vertex set $V_1 \cup V_2$. Then $G$ contains $\#(E)^2/\#(V_2)$ length two paths with both endpoints in $V_1$. $G$ also contains $\#(E)^4/\#(V_1) \#(V_2)$ length four cycles.
\end{theorem}
\begin{proof}
	s
\end{proof}

 \end{document}